\s{Differences between intuitionistic logic and classical logic}

If you'll remember,

\begin{description}
  \item[classical logic] is where things are true or false.
  \item[intuitionistic logic] is where things are provable or not provable.
\end{description}

It seems like a subtle difference, but there are some rather annoying
consequences, which I'll go through in this section.

\ss{Your English teacher lied to you}

In classical logic, a well-known rule is the double-negative:

\begin{textmath}
    For all $X$,

    \begin{equation}
        X \iff \lnot \parens{\lnot X}
    \end{equation}
\end{textmath}

This isn't true in intuitionist logic. Well, one half of it is true. Namely:

\begin{definition}[Double negation, right]
    For all $X$,

    \begin{equation}
        X \implies \lnot \parens{\lnot X}
    \end{equation}
\end{definition}

I proved this in \cref{expl:doubneg1}.

The other half can't be proven, though

\begin{textmath}
    For all $X$,

    \begin{equation}
        X \impliedby \lnot \parens{\lnot X}
    \end{equation}
\end{textmath}

Let's try to prove that:

\begin{proof}[Double negation, left]
    For all $x$

    \begin{eqnarray}{cl}
          & X \impliedby \lnot \parens{\lnot X} \\
        = & \lnot \parens{\lnot X} \implies X \\
        = & \lnot \parens{X \implies \bot} \implies X \\
        = & \parens{\parens{X \implies \bot} \implies \bot} \implies X \\
    \end{eqnarray}

    We can't Curry the type, because the operators are left-associative. It's
    not provable.
\end{proof}

\pg{Forall}

I keep having to write ``For all $x$''. Conveniently, mathematicians have a
symbol for this: $\forall$. In text, it's clearer and easier to type out ``for
all''. However, a lot of people will use the $\forall$ symbol when writing
things by hand.

\sss{Negating foralls}

Remember, this is not provable:

\begin{equation}
    \fall X \semic X \impliedby \lnot \parens{\lnot X}
\end{equation}

To the untrained eye, this would then mean:

\begin{eqnarray}{cl}
      & \lnot\parens{\fall X \semic X \impliedby \lnot \parens{\lnot X}} \\
    = & \fall X \semic X \notimpliedby \lnot \parens{\lnot X} \\
\end{eqnarray}

\xtb{But that's not the case!} There are certainly some cases where
$X \iff \lnot\lnot X$. But, there are also many cases where that doesn't
hold. So, you can't say that either version holds for all cases.

Instead, you just take off the $\fall$

\begin{textmath}
    For some $X$:

    \begin{equation}
        X \impliedby \lnot \parens{\lnot X} 
    \end{equation}
\end{textmath}

Likewise, 

\begin{textmath}
    For some $X$:

    \begin{equation}
        X \notimpliedby \lnot \parens{\lnot X} 
    \end{equation}
\end{textmath}

The thing to take away from all this is, double negation doesn't always hold in
intuitionist logic.

\ss{The law of excluded middle}

You might be asking ``well, since we can't prove left-double-negation, doesn't
that mean that left-double-negation would imply $\bot$?''.

You're probably not asking that, because that sounds like complete nonsense.

Since we can't prove

\begin{equation}
    \fall X \semic X \impliedby \lnot \parens{\lnot X}
\end{equation}

Wouldn't that therefore mean that

\begin{equation}
    \fall X \semic \parens{X \impliedby \lnot \parens{\lnot X}} \implies \bot
\end{equation}

Well, no. You would be making another assumption from classical logic, called
\term{the law of excluded middle}, or \term{decidability}. That is:

\begin{equation}
    \label{eq:lem}
    \fall X \semic X \tor \lnot X
\end{equation}

where $X$ is a proposition.

Basically, given some statement $X$,

\begin{description}
  \item[In classical logic] you can always \term{decide} whether or not $X$ is
    true.
  \item[In intuitionistic logic] that's not the case. There are values of $X$
    that are \term{undecideable}.
\end{description}

Remember the liar's paradox:

\begin{textmath}
    This statement is false.
\end{textmath}

That statement cannot be said to be true or false. It's \term{undecidable}.

Remember, to negate a statement starting with ``for all'', you just need to find a
case where the statement doesn't hold. In the case of the decidability law, I've
just found a case where the decidability law doesn't hold, so therefore it can't
hold for all propositions.

\ss{De Morgan's laws}

The undecidability law and the lack of left-double-negation are really the core
things you need to worry about with intuitionist logic. The next thing to look
out for are De Morgan's laws.

\begin{definition}[De Morgan's first law]
    \begin{equation}
        \label{eq:demorgan1}
        \lnot\parens{P \tand Q} \iff \lnot P \tor \lnot Q
    \end{equation}

    You can think of this like this:

    \begin{itemize}
      \item I told you that
        \begin{enumerate}
          \item I went to the park, and
          \item I went to the grocery store
        \end{enumerate}
      \item If you can prove either
        \begin{enumerate}
          \item I didn't go to the park, or
          \item I didn't go to the grocery store
        \end{enumerate}

        then I lied to you.
    \end{itemize}
\end{definition}

\begin{definition}[De Morgan's second law]
    \begin{equation}
        \label{eq:demorgan2}
        \lnot\parens{P \tor Q} \iff \lnot P \tand \lnot Q
    \end{equation}

    This is the dual of the first law

    \begin{itemize}
      \item I told you that
        \begin{enumerate}
          \item I went to the park, or
          \item I went to the grocery store
        \end{enumerate}

      \item If you can prove that
        \begin{enumerate}
          \item I didn't go to the park, and
          \item I didn't go to the grocery store
        \end{enumerate}

        then I lied to you
    \end{itemize}
\end{definition}

Remember that each of those laws is really two laws. For instance, the first law
can be stated as:

\begin{textmath}
    \begin{equation}
        \lnot\parens{P \tand Q} \implies \lnot P \tor \lnot Q
    \end{equation}

    and

    \begin{equation}
        \lnot\parens{P \tand Q} \impliedby \lnot P \tor \lnot Q
    \end{equation}
\end{textmath}

I ``proved'', intuitively, the first branch of each law. Let's look at the
second branch of each law.

\begin{remark}[De Morgan's first law, part 2]
    \begin{equation}
        \lnot\parens{P \tand Q} \impliedby \lnot P \tor \lnot Q
    \end{equation}

    \begin{itemize}
      \item If I tell you
        \begin{enumerate}
          \item I didn't go to the grocery store, or
          \item I didn't go to the park
        \end{enumerate}

      \item If you can prove
        \begin{enumerate}
          \item I went to the grocery store, and
          \item I went to the park
        \end{enumerate}

        then I lied to you
    \end{itemize}
\end{remark}

\begin{remark}[De Morgan's second law, part 2]
    \begin{equation}
        \lnot\parens{P \tor Q} \impliedby \lnot P \tand \lnot Q
    \end{equation}

    \begin{itemize}
      \item I told you
        \begin{enumerate}
          \item I didn't go to the park, and
          \item I didn't go to the grocery store
        \end{enumerate}
      \item If you can prove
        \begin{enumerate}
          \item I went to the park, or
          \item I went to the grocery store
        \end{enumerate}

        then I lied to you.
    \end{itemize}
\end{remark}

So, all and all, there are four laws there. One of them isn't true. Can you
figure out which one?

I'll insert a page break so you can think about it. You know what, I'll make an
entire new section

