\s{Morphisms}

Let's review what I supposedly have taught you so far and how all the various
things tie in together.

\begin{enumerate}
  \item There are these things called \term{function}s, which take some input and spit
    out some output.
  \item There are these things called \term{type}s, which describe the ``type''
    of thing some object is.
  \item There is this weird curly brace notation for constructing types, called
    \term{comprehension notation}.
  \item We can curry and uncurry types to make function semantics slightly
    easier. Namely, it's a lot easier to play around with partial functions if
    the function type is curried.
  \item If a function has type $A \to B$, then $A$ is the domain and $B$ is the
    codomain.
  \item You can prove things by expressing propositions as types and proofs as
    functions.
\end{enumerate}

I'm going to introduce a new term: \term{isomorphic}. Some type $A$ is
\term{isomorphic} to $B$ if I can convert $A$ to $B$ to $A$ without any loss of
data.

\begin{definition}[Isomorphism]
    If there exist two functions $f$ and $g$ such that

    \begin{textmath}
        \begin{equation}
            f : A \to B
            g : B \to A
        \end{equation}

        and, for all $x : A$,

        \begin{equation}
            \eva{g}{\eva{f}{x}} = x
        \end{equation}

        then $f$ and $g$ form an \term{isomorphism}.
    \end{textmath}

    An isomorphism between two types if and only if two types are
    isomorphic. Thus, if you find an isomorphism, you can conclude that the two
    types are isomorphic. Likewise, if you are given that the two types are
    isomorphic, then you can conclude that there exists an isomorphism.
\end{definition}

\begin{theorem}
    
\end{theorem}
