\s{Differences (continued) (continued)}

Alright, that section was getting pretty long! You have probably guessed by now,
but the first part of the first De Morgan law is the one that isn't true in
intuitionistic logic. Why?

Well... I can't really think of a good intuitive reason. You can prove (and I
will) that De Morgan's first law only holds if you assume these two axioms:

\begin{enumerate}
  \item The law of excluded middle (LEM). That is,

    \begin{equation}
        \fall P \comma P \tor \lnot P
    \end{equation}

  \item Double negation

    \begin{equation}
        \fall P \comma P \iff \lnot \lnot P
    \end{equation}
\end{enumerate}

If you'll remember, those are the two axioms that separate intuitionistic logic
from classical logic. Anyway. Let's get going

\begin{theorem}[De Morgan's first law, part 1 (ish)]
    \begin{equation}
        \parens{
          \lnot\parens{P \tand Q} \implies \lnot P \tor \lnot Q
        }
        \iff
        \parens{
          \parens{
            \fall X \comma X \tor \lnot X
          }
          \tand
          \parens{
            \fall X \comma X \iff \lnot \lnot X
          }
        }
    \end{equation}

    Hmm. There are actually two theorems here that we have to prove!
    \begin{enumerate}
      \item You can prove LEM and double negation, given De Morgan's law 1.1
      \item You can prove De Morgan's law 1.1 given LEM and double negation
    \end{enumerate}

    The second seems considerably easier, so we'll start with that.

    \begin{proof}
        Here's what we're trying to prove:

        \begin{equation}
            \parens{
                \parens{
                    \fall X \comma X \tor \lnot X
                }
                \tand
                \parens{
                    \fall X \comma X \iff \lnot \lnot X
                }
            }
            \implies
            \parens{
                \lnot\parens{P \tand Q} \implies \lnot P \tor \lnot Q
            }
        \end{equation}

        We can do some Currying magic to produce

        \begin{equation}
            \parens{
                \fall X \comma X \tor \lnot X
            }
            \to
            \parens{
                \fall X \comma X \iff \lnot \lnot X
            }
            \to
            \lnot\parens{P \tand Q}
            \to
            \lnot P \tor \lnot Q
        \end{equation}

        That still looks incredibly confusing. Let's do the thing with the line

        \begin{alignmath}{rcl}
            a & : & \fall X \comma X \tor \lnot X \\
            b & : & \fall X \comma X \iff \lnot \lnot X \\
            c & : & \lnot\parens{P \tand Q} \\
            \tabucline \\
            \text{goal} & : & \lnot P \tor \lnot Q \\
        \end{alignmath}

        Let's unfold the definition of $\lnot$ to produce:

        \begin{alignmath}{rcl}
            a & : & \fall X \comma X \tor \parens{X \implies \bot} \\
            b & : & \fall X \comma X \iff \parens{\parens{X \implies \bot} \implies \bot} \\
            c & : & P \to Q \to \bot \\
            \tabucline \\
            \text{goal} & : & \parens{P \to \bot} \tor \parens{Q \to \bot} \\
        \end{alignmath}

    \end{proof}
\end{theorem}