\s{Combining not, and, and or}

These are some rules for combining $\lnot$, $\land$, and $\lor$ together. You
should verify these yourself (like with examples and stuff).

\begin{itemize}
\item $\lnot\parens{\lnot A} \iff A$
\item $\lnot\parens{A \land B} \iff \parens{\lnot A} \lor \parens{\lnot B}$
\end{itemize}

The last rule is

\[\lnot\parens{A \lor B} \iff \parens{\lnot A} \land \parens{\lnot B} \semicolon
\forall A,B\]

However, this can be derived from the last law. Let $C = \lnot A$, and
$D = \lnot B$. By making the aforementioned substitutions, we know already that

\[
\begin{array}{rcll}
\lnot\parens{\parens{\lnot C} \land \parens{\lnot D}} & \iff & C \lor D & \semicolon \forall C,D \\
C \lor D & \iff & \lnot\parens{\parens{\lnot C} \land \parens{\lnot D}} & \semicolon \forall C,D \\
\lnot\parens{C \lor D} & \iff & \lnot\parens{\lnot\parens{\parens{\lnot C} \land \parens{\lnot D}}} & \semicolon \forall C,D \\
\lnot\parens{C \lor D} & \iff & \parens{\lnot C} \land \parens{\lnot D} & \semicolon \forall C,D \\
\end{array}
\]

\begin{ExcList}
  \Exercise{Given $A$, is it always the case that $A \iff A$?}  \Answer{Yes. We
    know that $A \implies A$. Remember, $A \iff A$ is just saying that
    $\parens{A \implies A} \land \parens{A \impliedby A}$. Also recall that
    $A \implies B$ is the lazyman's way of writing $B \implies A$. Thus, if you
    know $A \implies A$, then it must be true that $A \implies A$, and therefore
    $A \iff A$. }

  \Exercise{Given $A$, is it the case that $A \land A \iff A$?}
  \Answer{Yes. There are two cases we need to deal with here:
    \[
    \left\{
      \begin{array}{rrcl}
        \true  \to & \true \land \true   & \iff & \true  \\
        \false \to & \false \land \false & \iff & \false \\
      \end{array}
    \right.
    \]
    In both cases, $A \land A \iff A$. Q.E.D. This technique is called ``proof
    by exhaustion''. We named every possible case --- in this case, there were
    only two --- and proved the theorem for each of them.  }

  \Exercise{Given $A$ and $B$, is it always the case that
    $A \land B \iff B \land A$?}

  \Answer{Yes. In the previous problem, we showed that $A \land A \iff A$. Thus,
    if $A$ and $B$ are both \truenm, or are both \falsenm, the answer is
    yes. Thus, the only case we need to consider is that in which $A$ is \truenm
    and $B$ is \falsenm.\footnote{You may be wondering why I'm not considering
      the case when $A$ is \falsenm, and $B$ is \truenm. However, as we showed
      earlier, it's the case that if $A \iff B$, then $B \iff A$. Thus if
      $A \land B \iff B \land A$, then $B \land A \iff A \land B$}

    If $\true \land \false \iff \false$, and $\false \land \true \iff \false$.}

  \Exercise{Given $A$, $B$, and $C$, is it always the case that $\parens{A \land B} \land
    C \iff A \land \parens{B \land C}$?}
  \Answer{Yes.}

  \Exercise{Given $A$, is it the case that $A \lor A \iff A$?}
  \Answer{Yes.}

  \Exercise{Given $A$ and $B$, which are both True/False values, is it always
    the case that $A \lor B \iff B \lor A$?}
  \Answer{Yes.}

  \Exercise{Given $A$, $B$, and $C$, is it always the case that $\parens{A \lor B} \lor
    C \iff A \lor \parens{B \lor C}$?}
  \Answer{Yes.}

  \Exercise{Given $A$, $B$, and $C$, what is the result of
    $A \land \parens{B \lor C}$?}
  \Answer{$A \land \parens{B \lor C} \iff \parens{A \land B} \lor \parens{A
      \land C}$}
\end{ExcList}