\s{Set construction}

Let's invent a set.

\begin{equation}
    Q \ce \mset{7, 7, 9, 5, 10, 1, 6, 6, 2, 10}
\end{equation}

There we go. Remember, order and multiplicity don't matter. But, for the sake of
clarity, let's put the elements in order, and deduplicate them.

\begin{equation}
    Q \ce \mset{1, 2, 5, 6, 7, 9, 10}
\end{equation}

Yay! You may have noticed that I slipped in the word \term{element} into the
previous sentence. Objects in the set are called \term{elements}. 

It's too strenuous on our weak mathematical hands to write ``$10$ is an element
of $Q$'', so instead we have the symbol $\in$. $\in$ is a very terrible attempt
at drawing an E. If you can't remember what $\in$ is, think ``{$\in$}lement
of''.

Let's look at some examples:

\begin{equation}
    Q \ce \mset{1, 2, 5, 6, 7, 9, 10}
\end{equation}
\begin{equation}
    6 \in Q
\end{equation}
\begin{equation}
    7 \in Q
\end{equation}
\begin{equation}
    11 \notin Q
\end{equation}

What if we want to say ``both $6$ and $2$ are elements of $Q$'', in one
expression? We could write it out like:

\begin{equation}
    \parens{6 \in Q} \land \parens{2 \in Q}
\end{equation}

But that's too cumbersome, so instead we'll write 

\begin{equation}
    2, 6 \in Q
\end{equation}

\pg{But won't that get confusing?}

Only if we put parentheses or braces around $2$ or $6$.

\begin{equation}
    \mset{2, 6} \in Q
\end{equation}

That's confusing, don't do that (yet).

\pg{The null set}

The null set is the simplest set, as it contains no elements. ``Null set'' takes
too long to write, so we use $\emptyset$ instead.

\begin{equation}
    \nil \ce \mset{}
\end{equation}

We can even have sets of other sets:

\begin{equation}
    \mset{\mset{1,2,3},\mset{4,5,6},\mset{3,1,6}}
\end{equation}

\ss{Notation and construction of infinite sets}

By now, you're hopefully familiar with the simple bracket notation for finite
sets:

\begin{equation}
    A \ce \mset{4,5,6,2,56,7573,22,20}
\end{equation}

It's not very hard. Next, we have the ``extrapolation notation'' of infinite
sets:

% \newcommand{\cdots}{\cdot\cdot\cdot}

\begin{rclmath}
    \N & \ce & \mset{0,1,2,3,4,5,6,7,8,9,10,11,12,\cdots} \\
    \Z & \ce & \mset{\cdots,-4,-3,-2,-1,0,1,2,3,4, \cdots} \\
\end{rclmath}

That notation is okay when there is a very obvious pattern. What about this:

\begin{equation}
    \Q \ce \mset{
        \frac{0}{1}
      , \frac{1}{1}
      , \frac{1}{-1}
      , \frac{0}{-1}
      , \frac{-1}{-1}
      , \frac{-1}{1}
      , \frac{-1}{2}
      , \frac{0}{2}
      , \frac{1}{2}
      , \frac{2}{2}
      , \frac{2}{1}
      , \frac{2}{-1}
      , \frac{2}{-2}
      , \frac{1}{-2}
      , \cdots
    }
\end{equation}

The pattern is not very obvious, and it requires a ton of repetition. For
instance

\begin{equation}
      \frac{0}{1}
    = \frac{0}{-1}
    = \frac{0}{2}
    = \frac{0}{-2}
\end{equation}

Yet, each of those is represented as a unique element in $\Q$. A much better
notation would be:

\begin{equation}
    \Q \ce \mset{\frac{x}{y} \semic x,y \in \Z \land y \neq 0}
\end{equation}

Okay, how do you read that?

\begin{iquote}
    \begin{equation}
        \Q \ce \mset{\frac{x}{y} \semic x,y \in \Z \land y \neq 0}
    \end{equation}

    $\Q$ is the set of all fractions $x$ over $y$, such that $x$ and $y$ are
    both in $\Z$, and $y$ is not equal to $0$.
\end{iquote}

That notation is sort of hairy, but you'll get used to it pretty quickly.

\ss{This statement is false}

Let's think about this situation:

\begin{textmath}
    There's a barber in a small town. For each man in the town, the barber cuts
    the man's hair iff the man in question doesn't cut his own hair.
\end{textmath}

Let's put this another way

\begin{textmath}
    For some man \xtb{Man} in the town:

    \begin{alignmath}{rl}
             & \text{Barber cuts \xtb{Man}'s hair} \\
        \iff & \text{\xtb{Man} doesn't cut his own hair} \\
    \end{alignmath}
\end{textmath}

Okay, that's simple enough.

\begin{textmath}
    Does the barber cut his own hair?
\end{textmath}

Okay, wait. Hmm. Well. Let's plug $\mathbf{Barber}$ into that definition above.

\begin{alignmath}{rl}
         & \text{Barber cuts \xtb{Barber}'s hair} \\
    \iff & \text{\xtb{Barber} doesn't cut his own hair} \\
\end{alignmath}

Well. Hmm. Crap! We've arrived at a contradiction.

\xtb{Why are you telling me about this?}

Well, that little situation I described just now is called \term{Russell's
  paradox}. A philosopher-stroke-mathematician named Bertrand Russell proposed
it in the early 20\textsuperscript{th} century, as a reason why we need a
formalization of set theory.

\xtb{I don't get how the barber thing applies to set theory}

Well, Russell actually proposed this:

\begin{textmath}
    Let $S$ be the set of all sets.

    \begin{equation}
        A \ce \mset{x \in S \semic x \notin x}
    \end{equation}
\end{textmath}

Or, in a slightly clearer notation

\begin{alignmath}{rl}
         & \text{\(s\) is in \(A\)} \\
    \iff & \text{\(s\) is not in \(s\)} \\
\end{alignmath}

Well, what about $A$?

\begin{alignmath}{rl}
         & \text{\(A\) is in \(A\)} \\
    \iff & \text{\(A\) is not in \(A\)} \\
\end{alignmath}

Crap.

\sss{Okay, so how do we fix this?}

Look at that comprehension notation:

\begin{equation}
    A = \mset{ x \in {P} \semic \eva{{f}}{{x}}}
\end{equation}

\begin{description}
    \item[$x$] describes what each element of $A$ will look like
    \item[$P$] is a set from which every element of $A$ is drawn
    \item[$\eva{f}{x}$] is a condition about $x$ that has to hold true if it
    is to be included in $A$.
\end{description}

This is fine. If we stick to this, then we're golden. The problem comes when you
don't include $P$ in the equation.

\begin{equation}
    A = \mset{ x \semic \eva{{f}}{{x}}}
\end{equation}

\begin{description}
    \item[$x$] describes what each element of $A$ will look like
    \item[$\eva{f}{x}$] is a condition about $x$ that has to hold true if it
    is to be included in $A$.
\end{description}

\xtb{But wait, I thought with Russell's paradox, all of the elements in $A$ were
  drawn from another set, $S$, which is the ``set of all sets''?}

Well, I lied, sort of. It turns out that the ``set of all sets'' doesn't
actually exist. Mathematicians in the early 20\textsuperscript{th} century
didn't know that. But, we know it now.

\ss{Review for the exercises}

Yes, there are exercises.

This section concerns constructing sets. There are three notations that we have
so far:

\begin{description}
  \item[Explicitly listing every element] This is for finite sets that are small
    enough to do so:

    \begin{rclmath}
        X & = & \mset{1,2,3,4,5} \\
        Y & = & \mset{1,2,4,7,10,12,15} \\
    \end{rclmath}
  \item[Extrapolation or interpolation] This is for finite or infinite sets that
    have a very obvious pattern.

    \begin{rclmath}
        \N & = & \mset{0,1,2,3,4,\cdots} \\
        \Z & = & \mset{\cdots,-2,-1,0,1,2,\cdots} \\
        E & = & \mset{0,2,4,6,8,10,\cdots} \\
        F & = & \mset{0,2,4,6,8,10,\cdots,438,440,442} \\
    \end{rclmath}

  \item[Comprehensions] This is that weird notation for describing finite or
    infinite sets with a less than obvious pattern

    \begin{rclmath}
        \Q & = & \mset{\frac{x}{y} \in \R \semic x,y \in \Z \land y \neq 0} \\ 
        E & = & \mset{x \in \N \semic \text{$x$ is even}} \\ 
        F & = & \mset{x \in \N \semic \text{$x$ is even} \land x \le 442} \\ 
    \end{rclmath}
\end{description}

\ss{Subsets and Supersets}

Remember when I said $\mset{2, 6} \in Q$, and we were really confused? In case
you don't remember, $Q = \mset{1, 2, 5, 6, 7, 9, 10}$. $\mset{2,6}$ is obviously
not an element of $Q$. However, $\mset{2, 6}$ is \xti{in} $Q$ but it's not an
element. It's weird. How do we express this notion?

The answer is with \term{subsets}. ``sub'' means ``smaller'', so a ``subset''
would be a ``smaller set''. $A$ is a subset of $B$ is all of the elements in $A$
are also in $B$. The notation is $A \subeq B$. Some people will read that as
``$A$ is contained in $B$''.

Referring to the previous example,

\[ \mset{2,6} \subeq Q \]

\pg{Wait, what is with the little line under the round half circley thing?}

So, actually, there are two types of subsets - \term{proper} and
\term{improper}. $A \subeq B$ is for improper subsets. $A \subof B$ is for
proper subsets. What's the difference, then?

$A \subeq B$ allows for the possibility that $A = B$. $A \subof B$ means that
$B$ is \xti{strictly larger} than $A$; there are some elements in $B$ that are
not in $A$.

\label{exists}
I've already defined $\forall$ in \cref{forall}. I'm now going to add another
symbol, $\exists$, which means ``exists''. I mention $\forall$, because
$\exists$ is used in the same context.

Anyway, back to business. I use $\subeq$ for improper subsets, and $\subof$ for
proper subsets. However, some people will use $\subof$ for improper subsets, and
something like $\subsetneq$ or $\subsetneqq$ for proper subsets. You have to
look out.

Here's something cool: $\nil$ has no elements, so it's a subset of every set.

\[ \nil \subeq A \sfall A \]

I'm sure you can figure this out, two sets are equal iff they have the same
elements. $A \subeq B$ means that $B$ has all of the elements that $A$ has.
$A \supeq B$ would mean that $A$ has all of the elements of $B$. So if both of
those are true, then $A = B$. That is:

\[ A = B \iff A \subeq B \land A \supeq B \]

So, what did we learn?

\begin{enumerate}
\item There are unordered collections with no multiplicity, called \term{sets}.
\item There are ordered collections with multiplicity, called \term{vectors}.
\item Given an object $o$ and a set $A$, you can ask $o \Qin A$.
\item You can pull smaller sets out of a set (I'll show you the mechanics
  below). The way to express that a set is ``embedded'' in another set, without
  being an element is with \term{subsets}. $A \Qsubeq B$ would be asking ``are
  all of the elements in $A$ also in $B$?''. (The converse doesn't necessarily
  have to be true).
\item There's a pretty easy set which has no elements, called $\nil$. An
  interesting property of $\nil$ is that it is a subset of all other sets.
\item Two sets are equal iff they have the same elements.
\end{enumerate}