\s{Elements}

Let's invent a set.

\[ Q = \mset{7, 7, 9, 5, 10, 1, 6, 6, 2, 10} \]

There we go. Remember, order and multiplicity don't matter. But, for the sake of
clarity, let's put the elements in order, and deduplicate them.

\[ Q = \mset{1, 2, 5, 6, 7, 9, 10} \]

Yay! You may have noticed that I slipped in the word \term{element} into the
previous sentence. Objects in the set are called \term{elements}. Yay, we
figured out what that word means!

It's too strenuous on our weak mathematical hands to write ``$10$ is an element
of $Q$'', so instead we have the symbol $\in$. $\in$ is a very terrible attempt
at drawing an E. If you can't remember what $\in$ is, think
``\raisebox{1pt}{$\in$}\hspace{0pt}lement of''.\footnote{You better think this,
  because it took me 30 minutes to get the alignment right. So, you know,
  remember $\in$ this way, or else\dots\label{30minalign}}

So, I'm going to ask you a question:

\[ 11 \Qin Q \]

(See, I used that thing from earlier with the question mark. I told you it would
help.) Well, the answer is no, $11$ is not an element of $Q$. As always,
mathematicians are too weak to write ``$11$ is not an element of $Q$'' every
time they want to say it, so instead they write

\[ 11 \notin Q \]

By contrast,

\[ 6 \in Q \]

What if we want to say ``both $6$ and $2$ are elements of $Q$''? Well, again, we
could write it out like:

\[ \parens{6 \in Q} \land \parens{2 \in Q} \]

But that's too cumbersome, so instead we'll write 

\[ 2, 6 \in Q \]

\pg{But won't that get confusing?}

Only if we put parentheses or braces around $2$ or $6$.

\[ \mset{2, 6} \in Q \]

That's confusing, don't do that (yet).

There's one more thing I need to go over, which is the null set - it's the
simplest set, as it contains no elements. ``Null set'' takes too long to write,
so we use $\emptyset$ instead.
