\s{Subsets and Supersets}

Remember when I said $\mset{2, 6} \in Q$, and we were really confused? In case
you don't remember, $Q = \mset{1, 2, 5, 6, 7, 9, 10}$. $\mset{2,6}$ is obviously
not an element of $Q$. However, $\mset{2, 6}$ is \xti{in} $Q$ but it's not an
element. It's weird. How do we express this notion?

The answer is with \term{subsets}. ``sub'' means ``smaller'', so a ``subset''
would be a ``smaller set''. $A$ is a subset of $B$ is all of the elements in $A$
are also in $B$. The notation is $A \subeq B$. Some people will read that as
``$A$ is contained in $B$''.

Referring to the previous example,

\[ \mset{2,6} \subeq Q \]

\pg{Wait, what is with the little line under the round half circley thing?}

So, actually, there are two types of subsets - \term{proper} and
\term{improper}. $A \subeq B$ is for improper subsets. $A \subof B$ is for
proper subsets. What's the difference, then?

$A \subeq B$ allows for the possibility that $A = B$. $A \subof B$ means that
$B$ is \xti{strictly larger} than $A$; there are some elements in $B$ that are
not in $A$.

\label{exists}
I've already defined $\forall$ in \cref{forall}. I'm now going to add another
symbol, $\exists$, which means ``exists''. I mention $\forall$, because
$\exists$ is used in the same context.

Anyway, back to business. I use $\subeq$ for improper subsets, and $\subof$ for
proper subsets. However, some people will use $\subof$ for improper subsets, and
something like $\subsetneq$ or $\subsetneqq$ for proper subsets. You have to
look out.

Here's something cool: $\nil$ has no elements, so it's a subset of every set.

\[ \nil \subeq A \sfall A \]

I'm sure you can figure this out, two sets are equal iff they have the same
elements. $A \subeq B$ means that $B$ has all of the elements that $A$ has.
$A \supeq B$ would mean that $A$ has all of the elements of $B$. So if both of
those are true, then $A = B$. That is:

\[ A = B \iff A \subeq B \land A \supeq B \]

So, what did we learn?

\begin{enumerate}
\item There are unordered collections with no multiplicity, called \term{sets}.
\item There are ordered collections with multiplicity, called \term{vectors}.
\item Given an object $o$ and a set $A$, you can ask $o \Qin A$.
\item You can pull smaller sets out of a set (I'll show you the mechanics
  below). The way to express that a set is ``embedded'' in another set, without
  being an element is with \term{subsets}. $A \Qsubeq B$ would be asking ``are
  all of the elements in $A$ also in $B$?''. (The converse doesn't necessarily
  have to be true).
\item There's a pretty easy set which has no elements, called $\nil$. An
  interesting property of $\nil$ is that it is a subset of all other sets.
\item Two sets are equal iff they have the same elements.
\end{enumerate}