\s{Subsets and Supersets}
\label{s:subsets}
\label{s:supersets}
\label{s:subsets-supersets}
\label{s:subs-supers}
\label{s:subs}
\label{s:supers}

Remember back to \cref{s:construction}. I invented a set $Q$

\begin{equation}
    Q = \mset{1,2,5,6,7,9,10}
\end{equation}

and then asserted that

\begin{equation}
    \mset{2,6} \notin Q
\end{equation}

and we were really confused?

Well, why? $\mset{2,6}$ is \xti{in} $Q$, but not in the way we've seen so far:
it's not an \term{element} of $Q$. How do we express this notion?

Well, all of the elements in $\mset{2,6}$ are also in $Q$. Let's invent some
notation for that:

\begin{equation}
    \mset{2,6} \subeq Q
\end{equation}

For convenience, we'll also define the inverted operator

\begin{equation}
    Q \supeq \mset{2,6}
\end{equation}

\xtb{Well, then, wouldn't it be the case that $Q \subeq Q$?}

Well, yeah. All of the elements in $Q$ are also in $Q$.

It's actually how equality of sets is formally defined:

\begin{alignmath}{rcl}
    P = Q & \iff & \begin{array}{rl}
                             & P \subeq Q \\
                       \land & P \supeq Q \\
                   \end{array} \\
\end{alignmath}

\xtb{Well, how do we formally define that weird half circle line thing?}

By, the way, if $P \subeq Q$, then you're supposed to say ``$P$ is a
\term{subset} of $Q$''. ``Sub'' means ``smaller'', so a ``subset'' would be a
``smaller set''.

Anyway, to answer your question, like this:

\begin{textmath}
    $P$ is a subset of $Q$:

    \begin{equation}
        P \subeq Q
    \end{equation}

    If (and only if), for all $x$,

    \begin{alignmath}{rl}
        & x \in P \\
        \implies & x \in Q \\
    \end{alignmath}
\end{textmath}

Likewise, 

\begin{textmath}
    $P$ is a \term{superset} of $Q$:

    \begin{equation}
        P \supeq Q
    \end{equation}

    If (and only if), for all $x$,

    \begin{alignmath}{rl}
        & x \in P \\
        \impliedby & x \in Q \\
    \end{alignmath}
\end{textmath}

(Remember, $\forall$ means ``for all'').

\ss{Wait, what is with the little line under the round half circley thing?}

So, actually, there are two types of subsets - \term{proper} and
\term{improper}.

\begin{description}
  \item[$A \subeq B$] is for \term{improper subsets}
  \item[$A \subof B$] is for \term{proper subsets}
\end{description}

\xtb{What's the difference?}

\begin{description}
  \item[$A \subeq B$] allows for the possibility that $A = B$.
  \item[$A \subof B$] means that $B$ is \xti{strictly larger} than $A$; there
    are some elements in $B$ that are not in $A$.
\end{description}

I use $\subeq$ for improper subsets, and $\subof$ for proper
subsets. \xtb{However}, some people will use $\subof$ for improper subsets, and
something like $\subsetneq$ or $\subsetneqq$ for proper subsets. You have to
look out.

\ss{The power set}

The power set of $A$, denoted $2^A$, is the set of all subsets of $A$.

\begin{rlmath}
    & x \in 2^A \\
    \iff & x \subeq A \\
\end{rlmath}
