\s{Categories}

So, I've been a bit less than honest with you. I've never outright said it, but
I've sort of implied that, if you have some condition $p$, you can form a set

\[ \mset{x \mid p} \]

Remember when I brought up $\Set$, and presented it as ``the set of all sets''?
Well, surprise, I was lying! So, it turns out that, if you have some condition,
you can't always form a set based on that condition.

Case in point:

\[ A = \mset{x \mid x \notin x}\]

This should be read ``$A$ is the set of all sets $x$ for which $x$ is not an
element of itself''. With regard to $x \Qin x$, there are two possibilities:

\begin{enumerate}
\item $x \in x$
\item $x \notin x$
\end{enumerate}

Okay, let's look at the entire set $A$. $A \Qin A$

Well, if $A \in A$, then it must be true that $A \notin A$, because $A$ is the
set of all sets for which $x \notin x$. However, if $A \notin A$, then
$A \in A$, because $A$ is the set of all $x$ such that $x \notin x$. Therefore,
if $A \notin A$, then $A \in A$. In essence:

\[ A \in A \iff A \notin A \]

We've found an instance in which our intuition of sets fails. A
philosopher-stroke-mathematician named Bertrand Russell was the first to point
this out, in the early 20\textsuperscript{th} century.\cite{jech-set-theory} The
conclusion to draw from this result is that we need a more formal definition of
sets, not the slimy intuitive definition we already had.
% 
% Let's write an appendix on Bertrand Russell.

This is called ``disproof by counterexample''. We were trying to prove that some
property was universally true. However, we found one instance in which the
property was false, so therefore it can't be universally true.

Like I said above, we need some formal axioms for sets. The way we'll do this
is, instead of calling $\Set$ the ``set of all sets'', we're instead going to
call it the ``\xti{category} of sets''. Ooh yay we used the section title! So,
what's a category?

\nocite{w-cats}

There isn't much in the way of intuition for categories. Basically, a category
is a ``structure,'' or a ``network''. There are objects in the structure, and
then mappings (think functions), called ``arrows'' or ``morphisms'' between the
objects.

Categories are interesting because they are so stupidly abstract that they can
be used to describe most ideas in math. There are a lot of stupid things that
can describe all of math, so why are cateories special? Because, as far as we
know, category theory isn't riddled with contradictions.

A few paragraphs ago, I showed you Russell's paradox: that thing with the set
being both an element of itself and not an element of
itself.\footnote{Hereafter, that will be called Schrödinger's
  set.}\footnote{Just kidding.} That example shows that something as seemingly
cut-and-dry as the notion of a set can be riddled with contradictions, and thus
be hard to reason about. Categories aren't super intuitive, but they are free of
contradictions, and, as we'll see, the theory about them actually has some
pretty interesting results.
