\s{Operators and Functions}

This section will go over the general idea of functions --- and therefore
operators. Remember back to \cref{booleans}, I yammered on about the idea where
you could take a Boolean value, call it $A$, and then you could do $\lnot A$,
which would still be a Boolean value. Well, most people would call $\lnot$ an
``operator.'' Basically, it takes a Boolean, and gives you another Boolean. I'm
lazy, so instead of ``Boolean value'', I'm just going to write ``Boolean,'' or
sometimes ``Bool.''

\newcommand{\bool}{\mathbb{B}}
\newcommand{\boolnm}{$\mathrm{\bool}$}

It's convenient for us lazy math people to have a notation for ``$\lnot$ takes a
Boolean, and gives you another Boolean.'' First, we need a symbol for ``this is
a Boolean value.'' $\bool$. There we go. So, here's the notation for ``$\lnot$
takes in a Boolean and spits out a Boolean.''

\[ \lnot : \bool \to \bool \]

This property is called \xti{algebra}, in the sense that the operator $\lnot$
is \xti{algebraic}. An operator is \xti{algebraic} if it takes an element from
a given set, and then spits back out an element of that set.

So, what really is \boolnm? Well, it's the set of Booleans.

\xti{Wait, we're using sets already for stuff?}

Yes, silly, why do you think I was talking about them?

So, basically:

\[ \bool = \mset{\true, \false} \]

Here's something super cool and fun to do in math, that we'll do a lot. We're
going to take concepts that seem impossible to define, and define them. I bet
the average mathematician is capable of defining irony. It seems boring, but
it's actually pretty fun.

So, try to write down a definition of $\lnot$.

You probably can't do it, can you? It just seems so obviously true. Well, the
easiest thing to do is to just list some properties of $\lnot$, and then usually
a definition emerges.

We already know


What's the result of $\lnot\true$? Well, \falsenm, of course! So:

\[ \lnot\true = \false \]

We also know that $\lnot\parens{\lnot A} = A \sfall A$, so we can write down:

\[ \lnot \parens{\lnot A} = A \sfall A\]

(you should actually be writing this down on paper, by the way)

So, we're lazy, and we don't like writing $\sfall A$ or $\sfall A,B,C$ or
whatever every time we write something down. It just gets annoying. Instead,
we're going to use the symbol $\equiv$. You can read $\equiv$ as ``identically
equal to.'' So, $A \equiv B$ is just the lazyman's way of writing
$A = B \sfall A,B$. Thus, we can restate the previous definition as:

\[ \lnot \parens{\lnot A} \equiv A \]

Let's recollect all the information we have so far:

\begin{displaymath}
  \begin{array}{l}
    \bool                 =      \mset{\true,\false} \\
    \lnot                 :      \bool \to \bool \\
    \lnot\true            =      \false \\
    \lnot\parens{\lnot A} \equiv A \\
  \end{array}
\end{displaymath}

This is almost enough. We need one more thing. You could state that
$\lnot\false = \true$, which is \truenm, but not as useful as what we're going
to say:

\[ \parens{A = B} \implies \parens{\lnot A = \lnot B} \sfall A,B\]

You might be asking ``he used a $\forall$ and $=$, why didn't he use the cool
$\equiv$ symbol?'' Really, because it would have been wrong. $\equiv$ is really
a lazyman's amalgamation of $=$ and $\forall$. So, in that last thing, I was
talking about a situation involving three separate operators: the $=$ on the
left, the $\implies$, and the $=$ on the right. It was important to note that
the $\forall A,B$ applies to all of those operators collectively. $\equiv$ would
mean that the $\forall$ only applies to the $=$ operator.

Anyway, here's what we know:

\begin{displaymath}
  \begin{array}{l}
    \bool                 =        \mset{\true,\false} \\
    \lnot                 :        \bool \to \bool \\
    \lnot\true            =        \false \\
    \lnot\parens{\lnot A} \equiv   A \\
    \parens{A = B}        \implies \parens{\lnot A = \lnot B} \sfall A,B \\
  \end{array}
\end{displaymath}

This is enough to define $\bool$. Here's one quick exercise, then we'll move on.

\begin{ExcList}
  \Exercise{Using only the definition above, (and the definition of $=$), show
    that $\lnot False = True$.}  \Answer{
    \begin{proof}
      From the definition of $\lnot$:
      \[\lnot\true = \false\]
      By the commutativity of equality:
      \[ \false = \lnot \true \]
      From the definition of $\lnot$:
      \[\parens{A = B} \implies \parens{\lnot A = \lnot B} \sfall A,B\]
      Therefore,
      \[ \lnot \false = \lnot\parens{\lnot \true} \]
      From the definition of $\lnot$:
      \[\lnot\parens{\lnot A} \equiv A\]
      Therefore,
      \[\lnot\parens{\lnot \true} = \true \]
      By the transition of equality,
      \[ \lnot \false = \true \]
    \end{proof}
  }

  \Exercise{Use the following definition of $\land$ to prove:

    \begin{itemize}
    \item $A \land A \equiv A$
    \item $\parens{A \land B} \land C \equiv A \land \parens{B \land C}$
    \end{itemize}

    \begin{displaymath}
      \begin{array}{l}
        \land : \bool^2 \to \bool \\
        A \land \true \equiv A \\
        A \land B \equiv B \land A \\
      \end{array}
    \end{displaymath}

    {\footnotesize By the way, the $\bool^2$ thing just means it needs two
      elements of $\bool$ to do its job.}  }

  \Exercise{Do the same thing for $\lor$}

  \Exercise{Use this definition (and those aforementioned):
      \begin{displaymath}
        \lnot\parens{A \land B} \equiv \lnot A \lor \lnot B
      \end{displaymath}

      to prove

      \begin{displaymath}
        \lnot\parens{A \lor B} \equiv \lnot A \land \lnot B
      \end{displaymath}
  }

\end{ExcList}


\ss{Moving on}

Before we got distracted with all that proof stuff, we were talking about
operators. $\lnot$ is rather narrow in scope. It deals with literally two
values. There are some operators that deal with more values. 

You remember numbers, right? Well, let's look at an operator, which takes a
number, and adds 5 to it. We're going to call that operator $f$. Here's how
we're going to write $f$:

\[ f = \ld{x} x + 5 \]

\newcommand{\bracketset}[1]{\brackets{ #1 }}
% \newcommand{\evalmultat}[2]{#1 \bracketset{#2}}
\newcommand{\evalat}[2]{#1 _{\,\bracketset{#2}}}
\let\evalmultat\evalat

If we want to say $f$ evaluated where $\mlist{x = 1}$, we'll write

\[ \evalat{f}{x = 1} \]

Let's carry this out:

\begin{displaymath}
  \begin{array}{rcl}
    f & = & \ld{x} x + 5 \\
    \evalat{f}{x = 1} & = & \ld{x= 1} x + 5 \\
    & = & 1 + 5 \\
    & = & 6 \\
  \end{array}
\end{displaymath}

Hopefully you get the idea. $f$ evaluated where $x = something$ basically means
``replace $x$ with $something$, and carry out the calculation.''

That thing where  you have to wrap the parentheses around both of the things:

I'm going to write another operator, $g$, which takes two numbers. It adds $5$
to the first number, and subtracts $2$ from the second. Then it multiplies the
intermediate results together.

\[ g = \ld{x,y} \parens{x + 5} \times \parens{y - 2} \]

So, let's see what happens when we carry out $g$ when $x = 4$ and $y = 6$. First
of all, how do you write that?

\newcommand{\gfun}{\ld{x,y} \parens{x+5} \times \parens{y-2}}
\newcommand{\gfunevald}{\ld{x=4, y=6} \parens{x+5} \times \parens{y-2}}
\begin{displaymath}
  \begin{array}{rcl}
    \evalmultat{g}{x = 4, y = 6} & = & \gfunevald \\
                     & = & \parens{4 + 5} \times \parens{6 - 2} \\
                     & = & 9 \times 4 \\
                     & = & 36 \\
  \end{array}
\end{displaymath}

What are the signatures\footnote{(the thing with the colon and the arrows)} for
these functions?

Well, $f$ takes a number, and gives back a number. Our fancy-schmancy symbol for
``numbers'' is $\N$.

\[ f : \N \to \N \]

Thus, because $g$ requires 2 numbers to do its job, its signature is:

\[ g : \N^2 \to \N \]

\ss{Currying}

This is a nice point to break and talk about Currying.

Currying is basically another instance of mathematicians being clever and lazy
at the same time.

Remember $g$, right?

\[ g = \gfun \]

Well, what happens when we let $x = 4$, but leave $y$ alone?

\newcommand{\gfunprtevald}{\ld{x=4, y} \parens{x+5} \times \parens{y-2}}

\begin{displaymath}
  \begin{array}{rcl}
    \evalat{g}{x = 4} & = & \gfunprtevald \\
                      & = & \ld{y} \parens{4 + 5} \times \parens{y - 2} \\
                      & = & \ld{y} 9 \times \parens{y - 2} \\
  \end{array}
\end{displaymath}

So, in this case, $g$ took in a number, and gave us back a function with the
signature $\N \to \N$. So, in this case, $g$'s signature is actually:

\[ g : \N \to \parens{\N \to \N}\]

Thus it might be more semantic to write $g$ as:

\[ g = \ld{x} \parens{\ld{y} \parens{x + 5} \times \parens{y - 2}} \]

\xti{But wait, isn't that notation rather constrictive? What if we want to do
  $\evalat{g}{y = 3}$?}

Since it's pretty clear that $\to$ is right-associative --- i.e. it doesn't make
sense to write
$\parens{\ld{x} \lambda\mset{y}} \to \parens{x + 5} \times \parens{y - 2}$ ---
then we can be lazy and write:

\begin{displaymath}
  \begin{array}{l}
    g : \N \to \N \to \N  \\
    g = \ld{x} \ld{y} \parens{x + 5} \times \parens{y - 2} \\
    g = \ld{x, y} \parens{x + 5} \times \parens{y - 2} \\
  \end{array}
\end{displaymath}

And thus, it no longer seems wrong to write $\evalat{g}{y = 3}$.

\xtb{What's that $\lambda$ symbol?}

It's the Greek letter lambda, pronounced ``lamb-duh.''

\ss{Moving on Again}

So, we've --- in very little depth --- covered the idea of what a function looks
like. What about operators?

Well, operators are actually just really sneaky functions. Remember, the
definition for $\lnot$ looked an awful lot like our function.

\begin{displaymath}
  \begin{array}{l}
    \bool                 =        \mset{\true,\false} \\
    \lnot                 :        \bool \to \bool \\
    \lnot\true            =        \false \\
    \lnot\parens{\lnot A} \equiv   A \\
    \parens{A = B}        \implies \parens{\lnot A = \lnot B} \sfall A,B \\
  \end{array}
\end{displaymath}

However, it is really awkward to write unary operators (things that take only
one argument) in $\lambda$-notation. So, we just don't. Especially when the
definitions are recursive ($\lnot$'s definition uses $\lnot$ in the
definition). Your English teacher probably told you that it's not okay to use a
word in its own definition. Well, in math, it is okay.

\ss{Function Composition}

Okay, just two more little subsections, then you get to do all the wonderful
exercises! Yay!

Function composition is another example of mathematicians being lazy. Let's look
at that definition of $f$ from above:


\[ f = \ld{x} x + 5 \]

\newenvironment{alignedmath}{\begin{displaymath}\begin{array}{l}}{\end{array}\end{displaymath}}
\newcommand{\lto}{\lambda\to}

\begin{alignedmath}
  f : \N \to \N \\
  f = \ld{x}{x + 5}
\end{alignedmath}

In that case, because the function only does one thing (adds 5), the whole thing
about explaining that there's an $x$ there is sort of obvious

\[ f = \lto \parens{+ 5} \]

And even the $\lto$ is getting in the way. It's pretty obvious that
$\parens{+ 5}$ isn't a number, especially when we annotate $f$ with its
signature. So, we can be lazy, and write:

\begin{alignedmath}
  f : \N \to \N \\
  f = \parens{+ 5}
\end{alignedmath}

Yay for laziness!

Now, what if we want to add 5 to the number and then multiply the intermediate
result by $10$? Well, the verboseman's way of writing this is:

\begin{alignedmath}
  f : \N \to \N \\
  f = \ld{x} \parens{x + 5} \times 10
\end{alignedmath}

Well, let's see. There's probably an easier way to write that. First, we know
that $\times$ is commutative (we'll prove this later, don't worry), so we can write

\[f = \ld{x}  10 \times \parens{x + 5}  \]

Let's give the intermediate result a value:

\begin{alignedmath}
  f = \ld {x}  10 \times \parens{\ld {z = x} z + 5}
\end{alignedmath}

Basically $f$ says ``do $q$ first, then multiply the result of $q$ by $10$.''
Well, this seems a little verbose. We're sort of smushing the functions
$\parens{\times 10}$ and $\parens{+ 5}$ together. Well, luckily there's a way to
write this:

\begin{displaymath}
  f = \parens{\times 10} \circ \parens{+ 5}
\end{displaymath}

Note that order is important. When evaluating $\evalat{f}{\ldots}$, you do the
thing on the right-hand-side of $\circ$ first, then the thing on the left
afterwards.

You can read $\circ$ as ``compose'' or ``of'', depending on your taste.

% \ss{Formal definitions}

% Again, now that I've given you the intuition, let's go 