\ch{Sets, unary operators, and binary operators.}

This is a chapter on sets. What is a set, you ask? A set is a bunch of
things. Well, duh, right? With math, when we use a colloquial term, like ``set''
or ``group,'' we {\it usually} mean that the math term is similar in {\it
  concept} to the colloquial term, but we usually attach some strings.

The intuition for sets should hopefully be reasonably obvious. But, in the
interest of establishing a convention, I'll first cover the {\it intuition} (or
motivation) behind set theory in \cref{s: Intuition of a Set}, and then go over
the fine-print stuff in \cref{s: Definition of a Set}.

\s{Intuition of a Set}

This is what math people call ``intuitive set theory.'' Or rather, that's the
phrase we use when describing it to other people. As you will come to learn,
mathematicians are, on average, very pompous and pretentious. For that reason,
when us mathematicians talk about this version of set theory among ourselves, we
use the term ``naïve set theory.''\cite{w-naive-set-theory}

This is a quote from the Wikipedia article on naïve set theory:

\begin{quote}
  A set is a well defined collection of distinct objects. The objects that make
  up a set (also known as the elements or members of a set) can be anything:
  numbers, people, letters of the alphabet, other sets, and so on.
\end{quote}

That's a pretty good definition, right there. You might be asking what exactly
the term {\it well-defined} means. And if so, congratulations! That said, the
term {\it well-defined} isn't really well-defined.

Now it turns out that this notion of a set doesn't exactly work out, which is
why we have actual set theory, which I'll go over in \cref{ch: Definition of a
  Set}. But, for the time being, let's roll with this. 

\pg{Notation}

Here's how I'm going to distinguish ``set'' vs ``not a set:''

\begin{enumerate}
\item $0$ is not a set, it is instead the number $0$.
\item $468$ is not a set, it is instead the number $468$.
\item $\{0\}$ is a set, containing only the number $0$.
\item $\{468\}$ is a set, containing only the number $468$.
\item $\{0,1,2,3,4,5,6,7,8,9\}$ is a set, containing the numbers from $0$ to $9$.
\item $\{299,300,301,302,303,304,305,306\}$ is a set, containing the numbers
  from $299$ to $306$.
\item $\nil$ is a set, but doesn't have anything in it.
\end{enumerate}

You can also have sets of sets, and even combine those sets together to get new
sets. There are many ways to do that. The two most common of which are
``unions'' and ``intersections.''

\newcommand{\sfse}{\{468\}}                 %sfse = set four six eight
\newcommand{\sztn}{\{0,1,2,3,4,5,6,7,8,9\}} %sztn = set zero through nine
\newcommand{\sotn}{\{1,2,3,4,5,6,7,8,9\}}   %sotn = set one through nine

\begin{enumerate}
\item
  $\{ \{468\}, \{0,1,2,3,4,5,6,7,8,9\}, \{299,300,301,302,303,304,305,306\} \}$
  is a meta-set (a set of sets), containing the three aforementioned sets, in
  order. Note that by ``in order,'' I mean only that I listed the elements in
  the order they appeared before. Sets do not have any sort of intrinsic order.
  % Thank you, Benzrf
\item $\{468\} \cup \{0,1,2,3,4,5,6,7,8,9\}$ is the ``union'' of $\sfse$ and
  $\sztn$. You can think of the union as ``smushing the two sets together to get
  a new set.'' This union comes out to be $\{468, 0,1,2,3,4,5,6,7,8,9\}$. You
  can read $A \cup B$ as ``$A$ cup $B$.''
\item $\sotn \cap \sztn$ is the ``intersection'' of $\sotn$ and $\sztn$ and can
  be thought of as finding the common elements. You should read $A \cap B$ as
  ``$A$ cap $B$.'' This intersect comes out to be $\sotn$.
\end{enumerate}

So, what's the difference between the two?

Well, the intersect finds the elements that are in {\it both} sets. The union
finds the elements that are in {\it either or both} of the sets. 

\newpage
\sss{Exercises}

The solutions are on the next page.

\pg{Unions}

\begin{enumerate}
\item Make up some arbitrary set, and we'll call it $A$. $A$ can be any
  set. Explain why $A \cup \nil$ always turns out to be $A$, no matter what
  $A$ actually is.
\item Do the same thing as in the previous problem --- but --- this time,
  explain why $\nil \cup A$ always turns out to be $A$, no matter what $A$
  is. Note the difference between $\nil \cup A$ and $A \cup \nil$.
\item Examine your reasoning from the last 2 problems. Does your reasoning imply
  that $A \cup A$ is always $A$, no matter what $A$ actually is? Why or why not?
\item Does this mean that $\nil \cup \nil$ is always $A$, no matter what $A$ is?
  Why or why not?
\end{enumerate}

\newpage
\sss{Solutions}

Our apologies that these solutions are written in formalish math language. Our
native tongue is unfortunately fancy math language, and we have yet to translate
this.

\begin{enumerate}
\item $A \cup \nil$ is all of the items that are in either $A$ or
  $\nil$. Because there are no items in $\nil$, this would just be the items in
  $A$. The set of items in $A$ is just $A$. Thus, $A \cup \nil$ will always be
  $A$, no matter what $A$ is.
\item $\nil \cup A$ is all of the items that are in either $\nil$ or
  $A$. Because there are no items in $\nil$, this would just be the items in
  $A$. The set of items in $A$ is just $A$. Thus, $\nil \cup A$ will always be
  $A$, no matter what $A$ is.
\item $A \cup A$ is all of the elements in $A$ and $A$. Because those two sets
  are the same set, they therefore have the same elements. Thus, the set of
  elements in $A$ and $A$ is just the set of elements in $A$. The set of
  elements in $A$ is just $A$. Therefore, for any $A$, $A \cup A$ is always $A$.
\item In the previous examples, we showed that $A \cup A$ is always $A$ for any
  $A$. We could then say that $B \cup B$ is always $B$ for any $B$. If you
  decide that $B$ is $\nil$, then you can say $\nil \cup \nil$ is $\nil$. So,
  the assertion above --- $\nil \cup \nil$ is always $A$, no matter what $A$ is
  --- is false, because $\nil \cup \nil$ is always $\nil$, which means that the
  assertion fails if $A$ is not $\nil$.
\end{enumerate}

\newpage

\s{Definition of a Set}

% \item $\nil$ contains no elements. Hence, the union of both sets turns out to be $A$.
% \item $\nil$ contains no elements. Hence, as previously shown,
%   the union of both sets turns out to be $A$. The order of the sets does not matter.
% \item Yes. Both sets contain the same elements. Hence, the union of both sets turns out to be $A$.
% \item No. Both sets contain no elements. Hence, the union of both sets turns out to be $\nil$.
% \item Both turn out to be $\nil$. $\nil$ contains no elements. Hence, $A$ and $\nil$ share no elements.
% \item Yes. Both sets contain the same elements and must share all of their elements.