\ch{Sets}

So, I lied to you a little bit. Mathematicians don't usually use types. Some do,
but most don't. \term{Sets} are almost the same thing as types, and provide the
same abstraction capabilities. People tend to confuse sets and lists, so I
started with types, because the distinction is much clearer.

Informally, a set is a bunch of objects, with no particular order or notion of
duplication. That sounds like nonsense. In less nonsensical form:

\begin{equation}
    \mset{1,2,3} \text{ is a set.}
\end{equation}

Duplicate elements are not counted separately, so

\begin{equation}
    \mset{1,1,2,3}
\end{equation}

is the same set as $\mset{1,2,3}$. The order in which elements appear does not
matter, so 

\begin{equation}
    \mset{3,1,2,1,2}
\end{equation}

is also the same set.

Well, that's half true. Some sets are ordered, but every set does not have an
ordering.

Instead of the colon, the symbol $\in$ is used to indicate that some item is in
a set.

Remember the natural numbers, $\N$? If not,

\begin{equation}
    \N = \mset{0,1,2,3,4,5,\dots}
\end{equation}

All of arithmetic can be formalized with these 5 axioms. \cite{taylor-analysis, landau-analysis}

\begin{enumerate}
  \item $0$ is a natural number.
  \item For every natural number $n \in \N$, there is an immediate successor
    $\eva S n \in \N$.
  \item If $\eva S n = \eva S m$, then $m = n$. (In our previous terms, $S$ is
    an injection).
  \item There is no number $n \in \N$ such that $\eva S n = 0$
  \item This is a complicated expression:

    \begin{itemize}
      \item If there is a set $K$ such that
      \item $0 \in K$,
      \item the fact that $n \in K$ implies that $\eva S n \in K$,
      \item then $K$ contains every natural number.
    \end{itemize}
\end{enumerate}

The formalization of arithmetic from these axioms is incredibly boring, so I
won't cover it. If you really want to know about it, I recommend this book:
\fullcite{landau-analysis}.

The fifth axiom in that list sounds like nonsense, but it is actually incredibly
important. It sets up the basic idea of \term{mathematical induction}. The idea
is this:

\begin{itemize}
  \item Say you have some proposition, and the proposition is dependent on a
    number.
  \item For instance, $0 \le x$ is dependent on whatever $x$ is.
  \item If you know the proposition is true for $0$, and
  \item The fact that the proposition is true for some natural number $n$,
    implies that the proposition is true for $n$'s successor $\eva S n$. That is

    \begin{equation}
        \eva P n \implies \eva{P}{\eva S n}
    \end{equation}
  \item Then the proposition is true for every natural number.
\end{itemize}

It's an incredibly important proof technique.

% 1. Null set
% 2. Infinite sets
% 3. Sets containing other sets.
% 4. Comprehensions
% 5. Set equality
% 6. equivalence relations
% 7. Unions, intersections, subtractions, complements
% 8. Power sets
% 9. Cardinality
% 10. Russell's paradox
% 11. ZF
% 12. Axiom of choice and Zorn's lemma
% 13. The notion of proper classes

% Next section

% Different things that sets can do

% 1. Monoids, semigroups, groups, magmas
% 2. Rings
% 3. Fields

% Next section

% Vector spaces + linear algebra

% Next section

% Calculus

\input{chapters/ch4/s1-formalization.tex}
