\ch{Philosophy and/or FAQ}

by \pharpend

This book is written with a certain philosophy in mind. Explaining my philosophy
will answer a number of questions I am often asked.

First, I'll start with the license. he license I chose for this book is the GNU
Free Documentation License (FDL). Again, ``free'' refers to freedom, not
price. The FDL is similar in spirit to another license, the Creative Commons
Attribution-ShareAlike License (CC-SA). CC-SA is much more popular than the FDL,
mostly because it is much more general (e.g. you could distribute a painting
under CC-SA, but not the FDL).  The CC-SA license and the FDL are both
``copyleft'' licenses, in that they require that derivative works be licensed
under the same license.

\xti{So, why did you go with the FDL instead of CC-SA?}

Simply put, the CC-SA license is too general to fit our purposes. The FDL is
specifically designed for reference texts, so it has a clause requiring that the
work be made available in source form. The CC-SA has no such requirement.

To go on about this, I need to define ``freedom'' in academic works. This is a
modified version of the \link{https://gnu.org/philosophy/free-sw.html}{GNU
  Project's definition of free software}. In my view, an academic work is free
if you have:

\begin{itemize}
\item The freedom to use the digital document as you wish, for any purpose
  (freedom 0).
\item The freedom to reproduce exact copies of the document in any
  medium. (e.g. print the book). (freedom 1)
\item The freedom to distribute and/or sell exact copies of the document to
  whomever you choose, in any medium. (freedom 2).
\item The freedom to modify your own copy of the book. Access to the source is a
  precondition for this. (freedom 3)
\item The freedom to reproduce modified copies of the document in any
  medium. (e.g. print the book). (freedom 4)
\item The freedom to distribute and/or sell exact copies of your modified
  version to others, in any medium (freedom 5).
\end{itemize}

To guarantee freedom for everyone, we unfortunately have to restrict freedom 3 a
little bit. You can't modify the book in such a way that would restrict others'
freedom. Such ways would include

\begin{itemize}
\item Implementing digital restrictions management, or DRM.
\item Removing the license.
\item Releasing your modified version under a different license.
\end{itemize}

I suppose someone who ran in a different clique would wonder why people put up
so much fuss about something as silly as a license. The license explains exactly
what the reader can and can't do with my work. When my objective is freedom,
allowing everyone the maximum quantity of freedom requires some copyright
trickery, hence the 10-page-long license.

You would think that this freedom would be implicit in any academic
work. Unfortunately, that's not the case. 

An extreme instance of this trope of freedom restriction was a wonderful company
named Myriad Genetics. Myriad Genetics attempted to patent human
genes. Fortunately, the United States Supreme Court struck down this class of
patents in a
\link{http://www.supremecourt.gov/opinions/12pdf/12-398_1b7d.pdf}{unanimous
  decision}. For those of you who aren't American, unanimous US Supreme Court
decisions are incredibly rare. Myriad Genetics is headquartered within walking
distance from my house, so this was big news in my area.

Anyway, the point of all this is, freedom is important, especially in academic
works. Part of the reason I wrote this book is that there are very few free
textbooks.

To put it another way, it is of no benefit for the work to be nonfree. If the
work is free, I, as a writer, benefit from people giving me feedback, and
improving upon my work. You, as a reader, benefit from the freedom. The only
people who don't benefit are distributors (e.g. a publisher). However, in the
age of the internet, the need for a for-profit publisher isn't exactly clear.

To be clear, it is perfectly okay for a publisher to publish this book, and to
attempt to profit off of it. However, the publisher wouldn't have the
traditional nonfree monopoly over the book, which might discourage a publisher.
