\ch{Philosophy and/or FAQ}

Hello, everyone!

There exist a certain few questions which people ask me on an almost weekly
basis. I imagine the frequency of these questions will grow larger as the book
grows more popular. So, I'll dispel them now:

First, I'll start with the license. I suppose someone who ran in a different
clique would wonder why people put up so much fuss about something as silly as a
license. Hopefully this next stanza or two will shed light on that. If you
happen to hang out on the internet a lot, you're most certainly familiar with
the Creative Commons (CC) licenses.

If you aren't, basically the CC licenses are licenses for creative works. The
licenses are all at least partially free (in the sense of freedom). There are
two, if I recall correctly, that are free licenses.

Anyway, I digress. The license I chose for this book is the GNU Free
Documentation License (FDL). Again, ``free'' refers to freedom, not price. The
FDL is similar in spirit to one of the CC licenses, namely CC-ShareAlike
(CC-SA). CC-SA is much more popular than the FDL, mostly because it is much more
general (e.g. you could distribute a painting under CC-SA, but not the FDL).

The CC-SA license and the FDL are both ``copyleft'' licenses, in that they
require that derivative works be licensed under the same license.

\xtb{So, why did I go with the FDL instead of CC-SA?}

Simply put, the CC-SA license is too general to fit our purposes. The FDL is
specifically designed for reference works, so it has a clause requiring that
derivative work be made available in source form. The CC-SA has no such
requirement.

% My choice of the FDL complements some opinions I have about the academic community
% 
% Etc.
% 
% blah blah stallman
% blah blah free software
% blah blah free information
% blah blah myriad
% blah blah freedom is important
% blah blah non-freedom doesn't benefit anyone
