\s{The community}

Despite the fact that I used ``I'' in the first part of the book, LYSA
is actually a community project, and many people participate in the
writing of this book.

If you want to talk to us, or to other math people, come see us in {\tt
\#lysa} on Freenode. If you don't know what IRC is, or you don't have a
client set up, you can connect through
\link{http://webchat.freenode.net/?channels=lysa}{Freenode's webchat}.

If you have any questions about LYSA (or math), feel free to ask in the
IRC channel ({\tt \#lysa} on FreeNode in case you forgot).

If you want to submit a correction, or have some issue, or want to add
some content, really anything having to do with the content of the book,
you can visit \link{https://github.com/learnyou/lysa}{our GitHub
page}. We also have a \link{http://learnyou.org}{website} and a
\link{https://lysa.reddit.com/}{community on Reddit}.

\ss{Freedom}

This book is open-source. That means a number of things:

\begin{enumerate}
\item You are free to use this book as you like.
\item You are free to modify the book as you wish. The source is
      available at \btrurl{https://github.com/learnyou/lysa}.
\item You are free to redistribute and/or sell exact copies of this book
      as you wish.
\item You are free to redistribute and/or sell modified copies of this
      book as you wish.
\end{enumerate}

If, for instance, you are a schoolteacher and want to use this for your
class, you are free to edit it to your liking and give the modified copy
to your students. The only string I attach is, you have to allow anyone
to whom you give the book do the same thing (i.e. they have to be free
to copy/modify/change your version). 