\s{Target audience}

The explanation of why programming is useful is a good segue into
discussing the target audience.

When I was first writing the book, I wrote it in an effort to strengthen
my own understanding. So, the target audience was me. The very first
versions of this book were about a abstractish branch of math called
commutative algebra. Later on, it seemed more fitting to abstractly go
over the basics of math. That's what the current version of the book
does.

That doesn't answer the question: who is the target audience? Well,
people who want to learn basic and intermediate algebra, and to learn
why it's so interesting.

I mentioned this in \cref{real-intro}. Most books (and people) treat
math as a tool you can use for calculations. I treat math as a language
you can use to express your ideas. That's the core difference.  This
book will hopefully give you an interest in math itself, rather than
just a cursory knowledge of it.

With that in mind, my book is going to approach the topics much
differently than other books on the same topic. I rely very much on
abstraction and intuition.

It's that you know all of the basics about arithmetic: you should know
how to add, subtract, multiply, divide, and exponentiate. You should
also be familiar with the following sets, although it's not really hard
to pick up:

\begin{enumerate}
\item $\N \ce \mset{0,1,2,3,4, \dots }$
\item $\Z \ce \mset{\dots, \ng{4}, \ng{3}, \ng{2}, \ng{1}, 0,1,2,3,4, \dots }$
\item The real numbers $\R$.
\item The rational numbers $\Q$.
\item The irrational numbers $\I \ce \R \bs \Q$.
\item The complex numbers $\C$.
\end{enumerate}

If not, you should look at
\cref{basic-arithmetic}. \Cref{basic-arithmetic} is more or less a
blatant ripoff of \cite{landau-analysis}. I would like to think that my
appendix is much less dry, and does a bit more in the explaining
department. Landau's wonderful book is very dry. It's just theorem after
theorem after theorem. It's very rigorous, but you'll fall asleep after
reading a page.
