\s{Conventions used throughout}

You don't actually have to read this section, but it would be useful.

\begin{enumerate}
\item \code{Things in monospace} are either code snippets or commands to
be run in a terminal. I have separate stylings for \terminal{terminal
commands} and \code{inline code snippets}. That said, they are separate
but equal, at least for the time being.
\item The \S\ symbol refers to a section. So \S\ 3.2 means ``chapter 3,
section 2''.
\item Even though most of the writers are American, I still use the
British convention of putting periods after quotation marks: ``like
this''. The British convention is less ambiguous. If you see the
American convention anywhere in this book, please report it.
\item I will often recommend software. However, I will not recommend any
non-libre software, or any software that costs money.
\item ``I'' refers to me. ``We'' refers to both me and you, the
reader. ``You'' refers to, you guessed it, the reader. It's the
convention in academia to use the so-called ``royal we'', such as ``we
subtract 2 from both sides of the equation to obtain the result \dots''.

Sometimes, we will accidentally use the royal we, out of habit. Crap, I
just did it there! See? It's very difficult to avoid. Like any of the
other conventions herein, if you see it broken, please report the error
to the authors. You can use the \bugtracker, or, if you don't want to
make a gratis GitHub account, you can \emailme.

\item Oh yeah, sometimes I'll use \code{monospace} in things like URLs
or emails for the sake of disambiguity.

\item Most of the authors use some version of Linux. Hence, when there
are instructions for computer things (such as installing Haskell), I'll
write instructions for Linux, because that's what I know. There are two
solutions here:

  \begin{enumerate}
  \item You could try out Linux (it's gratis, and it's easy).
  \item If you know how to do the thing on your OS, and there aren't
  instructions for your OS, you could write up instructions and add them
  to the book. If you don't know how to do that, you could bring it up
  in the \bugtracker\ or \emailme.
  \end{enumerate}

\item If you see some number as a superscript in the middle of text:
like this\footnote{Hey, you found me!}, then the number refers to a
footnote. If the superscript number is in the middle of math, it's
probably just math.

\item If there's some number in brackets, like this: \cite{lyah}, then
it's a citation. If you're reading this as a PDF, you can actually click
on the number, and your PDF reader will take you to the relevant
bibliography entry. Go ahead, check it out! I'll wait. You can do the
same thing for footnotes and URLs.\footnote{Well, clicking the URL will
open up your web browser, but you get the point}

\label{conventions}
\item If you see something \term{in italics}, it's usually a vocabulary
word. Often there will be a term with a section number next to it:
\termref{like this}{d-equality}, usually somewhere in
\cref{appendix-d}. \Cref{appendix-d} is a reference section, which has
theorems, vocabulary, identities, stuff like that. So, the reference to
a section in \cref{appendix-d} next to a term points to the relevant
section in the appendix. Like all of the other references --- citations,
footnotes, URLs --- you can click on the section title, and your PDF
reader will take you there.
\end{enumerate}