\chapter{Functions and types}
\label{ch:functions}

There are two basic tools in abstract math. These tools are general enough to
allow you to express any idea.\footnote{You can actually prove using math that
  these tools allow expression of any idea. It's called \term{Turing
    completeness}.} These tools are \term{functions} and \term{types}.

These two tools\dots really, this whole chapter isn't very interesting. If you
already know what functions and types are, I suggest that you skip to the next
chapter.

I do, however, recommend that you do the exercises. They aren't very tedious. I
made a point throughout this book of writing non-tedious, but nonetheless
challenging exercises.

\s{Types}

Types are easier to understand, so I'll start with them. Every mathematical
object has a type. $3$, for instance, is a number. In math, it's useful to think
``what type of thing is this?''. That's all a type is.

Mathematicians love to be lazy. Instead of writing out ``$x$ is of type $T$''
every time, they'll instead write:

\begin{equation}
    x : T
\end{equation}

So, $3$ is a number. How do you express this?

This is actually a somewhat tricky question, just because there are a whole mess
of different types of numbers.

\begin{itemize}
  \item There's ``natural numbers'', which are $\mset{0, 1, 2, 3, 4, 5,
      \dots}$.
  \item There's the ``integers'', which are
    \(\mset {\dots, -2, -1, 0, 1, 2, \dots}\).
  \item Then, there's the ``real numbers'', which are way too hard to
    explain.
  \item There's ``imaginary numbers'', which are the opposite of real
    numbers.
\end{itemize}

The list goes on and on. I'll explain what all of these number types are later.
For now, we'll stick to the simplest type of numbers: natural numbers. The
symbol for natural numbers is $\N$.

\begin{equation}
    3 : \N
\end{equation}

Types even have a type. The type is called $\Type$.

\input{2/2-functions.ltx}
\input{2/3-currying.ltx}