\ch{Sets}\label{ch:sets}

In math, it's often useful to consider \xti{collections} of objects. There are
basically two types of collections:

\begin{description}
  \item[sets], in which order and multiplicity don't matter
  \item[vectors], in which order and multiplicity do matter.
\end{description}

For instance, $\mset{3, 4}$ and $\mset{4,3}$ are the same \xti{set}

\begin{equation}
    \mset{3,4} = \mset{4,3}
\end{equation}

but $\mvec{3,4}$ and $\mvec{4,3}$ are different \xti{vectors}.

\begin{equation}
    \mvec{3,4} \neq \mvec{4,3}
\end{equation}

Likewise, $\mset{20,38}$ and $\mset{38,20,38,20,20,20,20,20}$ are the same
\xti{set}

\begin{equation}
    \mset{20,38} =\mset{38,20,38,20,20,20,20,20}
\end{equation}

You guessed it, $\mvec{20,38}$ and $\mvec{38,20,38,20,20,20,20,20}$ are
different \xti{vectors}.

\begin{equation}
    \mvec{20,38} \neq \mvec{38,20,38,20,20,20,20,20}
\end{equation}

Sets are --- at least ostensibly --- much more important. More importantly, they
are much easier to understand.

Sets were first studied to extent by Georg Cantor, a German mathematician, in
the second half of the nineteenth century. Back in his own day, the results
Cantor found by studying sets were considered so thoroughly bizarre that many of
his colleagues simply refused to believe that Cantor could be right.  In the
end, Cantor turned out to be right all along. His ideas can be found in any
introductory text on mathematics---including this one.

You probably figured it out from above: the notation is $\mset{\text{Braces}}$
for sets, and $\mvec{\text{Parentheses}}$ for vectors.

If you can't remember whether to use braces \{the curly things\}, or parentheses
(the round things), remember: \xti{a \xtb{brace} is used to \xtb{set} a broken
  bone.} I don't have a horrible pun having to do with parentheses and vectors,
and for that, I apologize.

\input{3/1-elements.ltx}
\input{3/2-subs-supers.ltx}

