\ch{Sets}

In math, it's often useful to consider \xti{collections} of objects. There are
basically two types of collections: \term{sets} and \term{vectors}. Sets are
unordered, and multiplicity doesn't matter. Vectors are ordered, and
multiplicity does matter.

For instance, $\mset{3, 4}$ and $\mset{4,3}$ are the same \xti{set}, but
$\mvec{3,4}$ and $\mvec{4,3}$ are different \xti{vectors}.

Likewise, $\mset{20,38}$ and $\mset{38,20,38,20,20,20,20,20}$ are the same
\xti{set}. You guessed it, $\mvec{20,38}$ and $\mvec{38,20,38,20,20,20,20,20}$
are different \xti{vectors}. Sets are, for basic stuff, much more
important. Also, they are much easier to understand.

Sets were first studied to extent by Georg Cantor, a German mathematician, in
the second half of the nineteenth century. Back in his own day, the results
Cantor found by studying sets were considered so thoroughly bizarre that many of
his colleagues simply refused to believe that Cantor could be right.  In the
end, Cantor turned out to be right all along. His ideas can be found in any
introductory text on mathematics---including this one.

You probably figured it out from above: the notation is $\mset{\text{Braces}}$
for sets, and $\mvec{\text{Parentheses}}$ for vectors.

If you can't remember whether to use braces \{the curly things\}, or parentheses
(the round things), remember: \xti{a \xtb{brace} is used to \xtb{set} a broken
  bone.} I don't have a horrible pun having to do with parentheses and vectors,
and for that, I apologize.

\s{Elements}

Let's invent a set.

\[ Q = \mset{7, 7, 9, 5, 10, 1, 6, 6, 2, 10} \]

There we go. Remember, order and multiplicity don't matter. But, for the sake of
clarity, let's put the elements in order, and deduplicate them.

\[ Q = \mset{1, 2, 5, 6, 7, 9, 10} \]

Yay! You may have noticed that I slipped in the word \term{element} into the
previous sentence. Objects in the set are called \term{elements}. Yay, we
figured out what that word means!

It's too strenuous on our weak mathematical hands to write ``$10$ is an element
of $Q$'', so instead we have the symbol $\in$. $\in$ is a very terrible attempt
at drawing an E. If you can't remember what $\in$ is, think
``\raisebox{1pt}{$\in$}\hspace{0pt}lement of''.\footnote{You better think this,
  because it took me 30 minutes to get the alignment right. So, you know,
  remember $\in$ this way, or else\dots}

So, I'm going to ask you a question:

\[ 11 \Qin Q \]

(See, I used that thing from earlier with the question mark. I told you it would
help.) Well, the answer is no, $11$ is not an element of $Q$. As always,
mathematicians are too weak to write ``$11$ is not an element of $Q$'' every
time they want to say it, so instead they write

\[ 11 \notin Q \]

By contrast,

\[ 6 \in Q \]

What if we want to say ``both $6$ and $2$ are elements of $Q$''? Well, again, we
could write it out like:

\[ \parens{6 \in Q} \land \parens{2 \in Q} \]

But that's too cumbersome, so instead we'll write 

\[ 2, 6 \in Q \]

\xtb{But won't that get confusing?}

Only if we put parentheses or braces around $2$ or $6$.

\[ \mset{2, 6} \in Q \]

That's confusing, don't do that (yet).

There's one more thing I need to go over, which is the null set - it's the
simplest set, as it contains no elements. ``Null set'' takes too long to write,
so we use $\emptyset$ instead.

\s{Subsets and Supersets}

Remember when I said $\mset{2, 6} \in Q$, and we were really confused? In case
you don't remember, $Q = \mset{1, 2, 5, 6, 7, 9, 10}$. $\mset{2,6}$ is obviously
not an element of $Q$. However, $\mset{2, 6}$ is \xti{in} $Q$ but it's not an
element. It's weird. How do we express this notion?

The answer is with \term{subsets}. ``Sub'' means ``smaller'', so a ``subset''
would be a ``smaller set''. $A$ is a subset of $B$ is all of the elements in $A$
are also in $B$. The notation is $A \subeq B$. Some people will read that as
``$A$ is contained in $B$''.

Referring to the previous example,

\[ \mset{2,6} \subeq Q \]

\xtb{Wait, what is with the little line under the round half circley thing?}

So, actually, there are two types of subsets - \term{proper} and
\term{improper}. $A \subeq B$ is for improper subsets. $A \subof B$ is for
proper subsets. What's the difference, then?

$A \subeq B$ allows for the possibility that $A = B$. $A \subof B$ means that
$B$ is \xti{strictly larger} than $A$; there are some elements in $B$ that are
not in $A$.

Before I go on, I'm going to define a couple of other symbols. We're constantly
saying ``this thing exists'' and ``for all'', so I'm going to add a couple of
symbols. $\exists$ is for ``there exists'' --- notice the backwards E, as in
``exists''. $\forall$ is for ``for all'' --- notice the upside-down A, as in
``all''.

So, now that you've got all of this down (hopefully), I'll give you some exercises. 

\begin{ExcList}
  \Exercise{$\Qexists A \semic A \subof \nil$}
  \Answer{
    \nilnm, by definition, has no elements. For there to exist a proper subset
    of \nilnm, there would need to be a set with fewer elements. There cannot be
    fewer than zero elements, hence there is no proper subset of \nilnm.
  }

  \Exercise{$\forall A \semic \nil \Qsubeq A$}
  \Answer{
    Yes. It's obvious that $\nil \subof A$ if $A \ne \nil$. The only thing left
    is, $\nil \Qsubeq \nil$. $\nil \subeq \nil$ allows for $\nil = \nil$, which
    is obviously true.
  }
  
\end{ExcList}

% \begin{asidebox}
%   $\in$ was invented by Georg Cantor. It was later revealed
% \end{asidebox}
% \input{3/1-elements-subsets.ltx}
% \input{3/2-operators-functions.ltx}
% \input{3/3-unions-intersections.ltx}
% \input{3/4-categories.ltx}