\ch{Proofs}

By now you should be fairly familiar with the notions of institutionistic logic,
sets, and functions: at least to the point where you could try to explain the
concepts to your friends. 

In the introduction, I mentioned that there are 3 concepts with which you need
to be relatively familiar in order to do anything interesting. The first two
were sets and functions, respectively. This chapter covers the third concept:
proofs.

So, in case you forgot, what is a proof? Basically it's where you gather a bunch
of known things, and string them together to learn something new.

The standard approach is to go over the \xti{Peano axioms}, which can be used to
prove everything about the real numbers. As instructive as that is, it's also
really boring. Most of math isn't boring. However, if you do want to go over the
Peano axioms with rigor, \cite{landau-analysis} does a wonderful job. For now,
I'll just define what all of the sets are:

\begin{enumerate}
\item $\N$ is the set of ``natural numbers''. That is,
  $\mset{1,2,3,4, \ldots}$.\footnote{Some people assert that the natural
    numbers start with $0$, not $1$. It doesn't matter a lot. }
\item $\Z$ is the set of ``integers''. So,
  $\mset{\ldots, \ng 4, \ng 3, \ng 2, \ng 1, 0, 1, 2, 3, 4}$.

\item $\Q$ is the set of ``rational numbers''. These are the numbers that are
  \xti{ratios} of other numbers. Specifically:

  \[ \Q = \mset{ \frac{x}{y} \mid \parens{x \in \Z} \land \parens{y \in \N}}\]

\item $\R$ is the set of ``real numbers''. You can write any number down and
  it's a real number. Examples:

  \begin{enumerate}
  \item 3151.5158196818196
  \item 516818641
  \item 231
  \item 7455468765.4198160685198196064864984514514451031784158
  \end{enumerate}

  These are all real numbers.

\end{enumerate}

Now, I've conveniently set these up so that each succeeding set is a strict
superset of the previous set. That is,

\[ \N \subof \Z \subof \Q \subof \R \]

\xtb{Why is $\R$ called the set of ``real numbers''? Why isn't it called the set
  of ``numbers''? If $\R$ is the set of real numbers, are there some numbers
  that aren't real? Preposterous, I say!}

Well, actually, yes, there are numbers that aren't real. We'll get to that
later, though. I actually lied, we are going to go through the Peano axioms. But
not in a boring way, or in a super rigorous pull-your-hair-out way. Just in a
way so that I can talk about non-real numbers, and you'll know what the hell I'm
talking about.

Okay, first of all, how do we define $\N$? Well, we know that $1$ is a natural
number. Yay! this is the first axiom.

\[ 1 \in \N \]

Now, we know for each number $x \in \N$, that the next number is in $\N$. For
example, for $1$, $2$ is the next number, and $2$ is a natural number. Likewise,
for $2$, there is a next number, and that number is $3$. I think you get the
point. If our number is $x$, we'll call the next number ``the successor of
$x$'', and denote it as $\succ{x}$. So, this axiom we just described can be
stated as:

\[ \forall x \in \N\comma \succ{x} \in \N \]

In this case, we would say that $\N$ is ``closed under $\mathrm{succ}$'', or
that ``$\mlist{\N, \mathrm{succ}}$ forms a closure''.


