\ch{More stuff about sets}

By now, you hopefully have some idea into the basic intuition behind sets and
functions. Moreover, you've proven some cool stuff about them -- for instance,
you proved that a function is invertible iff it is bijective.

That's kind of cool, right? It's much easier to verify that a function is
bijective than it is to find its inverse. So, right off the bat, you can see if
a function is invertible without trying to invert it. Despite what you may
think, a common problem in math is to find inverse functions.

Speaking of functions, I'm going to define a really simple function:

\begin{alignedmath}
  \id : a \to a \\
  \evalat{\id}{x} = x
\end{alignedmath}

That function is about as simple as functions get. It's not a very interesting
function, but it's handy when defining things.

This is a slightly less simple function, but nonetheless important

\begin{alignedmath}
  flip : \parens{a \to b \to c} \to b \to a \to c \\
  \eva{flip}{f, x, y} = \eva{f}{y,x}
\end{alignedmath}

It takes a function, $f$, which takes an $a$ and a $b$, and then returns another
function with the arguments flipped. You won't usually see
$\eva{flip}{\ld{x, y} x-y, 3, 5}$ floating around. Instead

\begin{alignmath}{rcl}
  f & : & p \to q \to r \\
  \eva{flip}{f} & : & q \to p \to r \\
\end{alignmath}

Here's an infix operator:

\begin{alignedmath}
  \circ : \parens{b \to c} \to \parens{a \to b} \to a \to c \\
  \eva{\parens{f \circ g}}{x} = \eva{f}{\eva{g}{x}}
\end{alignedmath}

