\s{Propositions}
\label{props}

The first thing you need to understand is the notion that ``if x is true, then y
is also true. But, if y is true, it's not necessarily true that x is false.'' As
always, mathematicians are too lazy to write this stuff out by hand, so they
have notation for it.

\begin{enumerate}
  \item $a \implies b$ means that ``$a$ implies $b$''. It doesn't
    necessarily mean that $b$ implies $a$. It means that if $a$ is true,
    then $b$ is also true.

    Here's an example:

    If someone is decapitated, then they will die. So,

    \[ \text{Decapitated} \implies \text{Dead} \]

    However, if someone is dead, it doesn't necessarily mean that they
    were decapitated. They could have been shot, or stabbed, or had a
    heart attack. There are endless possibilities.

  \item $a \impliedby b$ is the same as writing $b \implies a$. It's
    sometimes convenient to use $a \impliedby b$ instead.
    $a \impliedby b$ should be read ``$a$ is implied by $b$''.

  \item When I strike through some mathematical operator, like this:
    $\notimplies$, it means that you can semantically put ``not'' in
    front of whatever the operator says. So, $\notimplies$ means ``not
    implies''. That doesn't make much grammatical sense in English, so
    ``does not imply'' might be better. Nonetheless, you get the point.

    Moving on from the example above:

    \[ \text{Decapitated} \implies \text{Dead} \]

    If someone is decapitated, then they're also dead (at least within a
    few seconds). However, if someone is dead, it's not necessarily true
    that they were decapitated.

    \[ \text{Decapitated} \notimpliedby \text{Dead} \]

  \item If something is not true, then I'll put a $\lnot$ in front of
    it. So, if I want to say that $a$ is false, then I'll write
    $\lnot a$.

  \item If I want to pose a question, I could just ask the question. For
    instance, ``Is $\lnot\text{Decapitated}\impliedby\lnot\text{Dead}$
    true?''.

    However, that quickly becomes difficult, usually when there are
    multiple assertions in a mathematical expression, and you don't know
    which one I'm asking about. Moreover, since I use the same font for
    text and math, if I have both, it might be hard to tell which is
    math and which is text. So, to help with ambiguity, I'll put a ?
    over the operator I'm asking about:

    \[ \lnot \text{Decapitated} \Qimpliedby \lnot\text{Dead} \]

    See, that's much easier.

  \item Now, on to that question - Is ``not decapitated'' implied by
    ``not dead''. Well, let's think about it. If someone is not dead,
    then they couldn't have been decapitated, because if they were
    decapitated, then they would be dead. Therefore, if someone is not
    dead, then they weren't decapitated.

    That word jumble was probably confusing. Mathematicians don't like
    to be confused. I'll make it symbolic for you.

    \begin{textmath}
        If

        \begin{displaymath}
            \text{Decapitated} \implies \text{Dead}
        \end{displaymath}

        then

        \begin{displaymath}
            \lnot \text{Decapitated} \impliedby \lnot\text{Dead}
        \end{displaymath}
    \end{textmath}

    This is another rule about propositional logic:

    \begin{textmath}
        If

        \[ a \implies b \]

        then

        \[ \lnot b \implies \lnot a \]
    \end{textmath}
\end{enumerate}