\s{Booleans as a monoid}

Okay, first of all, what the hell does the section title mean?

Let's go over it:

\begin{enumerate}
  \item \term{Booleans} the values true and false. ``Boolean algebra''
    is what we are doing in this section. Dealing with Booleans.

    \xti{Why are you capitalizing the word Boolean?}

    Because the term is named after a guy named George Boole, who
    studied them to extent.

  \item A \term{monoid} is a set of things, where you can combine the
    individual things together. Additionally, there has to be a
    \term{monoidal identity}.

    \xti{That doesn't help me.}

    Well, read this section, and then you'll get it. Probably.
\end{enumerate}

\sss{Logical-and and logical-or}

Sometimes we need to combine two pieces of logic together. There are two
ways we can do this - logical-or and logical-and.

I put logical- in front of them, because the mathematical meaning is
slightly different than the colloquial meaning.

Mathematicians are lazy, so we don't like to write ``logical-and''
whenever we want to say it, so instead we use the symbol $\land$.

$A \land B$ is true if (and only if) both $A$ and $B$ are true. If one
of them is false, then the entire thing is false.

\begin{alignmath}{rclcl}
    \false & \land & \_      & \ce & \false \\
    \true  & \land & x       & \ce & x \\
\end{alignmath}

The $\_$ just means ``you can ignore whatever goes here''.

On the other side, we have logical-or. The symbol for logical-or is
$\lor$.  $A \lor B$ is true if either $A$ or $B$ is true, or if both of
them are true. You could think of logical-or as being equivalent to the
colloquial ``and/or''.

\begin{alignmath}{rclcl}
    \true   & \lor & \_       & \ce & \true \\
    \false  & \lor & x        & \ce & x \\
\end{alignmath}

If you're having trouble remembering which symbol is logical-and and
which one is logical-or, remember that the logical-and symbol ---
$\land$ --- looks vaguely like an A.

\pg{Logical-not}

If you'll remember from \cref{props}, $\lnot$ ``negates'' a value. So,

\begin{alignmath}{rlcl}
    \lnot & \parens{\lnot x} & \ce & x \\
    \lnot & \true            & \ce & \false
\end{alignmath}

These values, true and false, are called \term{Booleans}. They are named after a
mathematician named George Boole who studied them to
extent.\cite{w-boolean-algebra}

\ss{If (and only if)}

I'm going to introduce some new notation: the $\iff$ symbol.

\[ \parens{A \iff B} \iff \parens{A \implies B} \land \parens{A
  \impliedby B} \]

Okay, what the hell is that?

\begin{textmath}
    If

    \[ A \implies B \]

    logical-and

    \[ B \implies A \]

    then

    \[ A \iff B \]

    Likewise, if

    \[ A \iff B \]

    then

    \[ A \implies B \]

    logical-and

    \[ B \implies A \]
\end{textmath}

The symbol \nm\iff\ should be read as ``if (and only if)''. Sometimes I'll write
the word iff --- with two `f's --- that's the same as ``if (and only if)''.

The properties of booleans are really boring, and really obvious. If you really
want to see exhaustive proofs, see \cref{d-booleans}. The only tricky one is De
Morgan's law:

\begin{textmath}
    For all $A$ and $B$,

    \begin{equation}
        \label{eq:demorgan1}
        \lnot\parens{A \land B} = \lnot A \lor \lnot B
    \end{equation}
\end{textmath}

Sometimes, mathematicians, such as myself, will use the symbol $\forall$ in the
place of ``for all''. Additionally, the exercises use the symbol $\equiv$:

\begin{textmath}
    If, for all $x$ and $y$,

    \begin{equation}
        x = y
    \end{equation}

    then

    \begin{equation}
        x \equiv y
    \end{equation}
\end{textmath}

\begin{ExcList}
    \Exercise[label={demorgan2}]{ For all $A$ and $B$, is it the case that

      \begin{equation}\label{eq:demorgan2}
          \lnot\parens{A \lor B} \Qequiv \lnot A \land \lnot B
      \end{equation}
    }

  \Answer{
      Start with the first law
      \begin{equation}
          \label{eq:-}
          \lnot\parens{a \land b} \equiv \lnot a \lor \lnot b
      \end{equation}

      Let $p = \lnot a$, $q = \lnot b$. One of the rules of algebra is, if
      $x = y$, then you can substitute one in for the other.

      \begin{equation}
          p \lor q \equiv \lnot\parens{\lnot p \land \lnot q}
      \end{equation}
      Another rule is, if you have an equation, and you do something to one side
      of the =, you have to do the same thing to the other side. Here, we're
      going to apply $\lnot$ to both sides of $\equiv$

      \begin{equation}
          \label{eq:-}
          \lnot\parens{p \lor q} \equiv {\lnot p \land \lnot q}
      \end{equation}
  }

  \Exercise{
    Is it the case that

    \begin{equation}
       a \lor \parens{b \land c} \equiv \parens{a \lor b} \land \parens{b \lor c}
    \end{equation}
    }

  \Answer{
    Yes. Start with the first property

    \begin{equation}
      a \land \parens{b \lor c} \equiv \parens{a \land b} \lor \parens{b \land c}
    \end{equation}

    Apply $\lnot$ to both sides

    \begin{equation}
      \lnot\parens{a \land \parens{b \lor c}} \equiv \lnot\parens{\parens{a \land b} \lor \parens{b \land c}}
    \end{equation}

    Apply DeMorgan's laws

    \begin{equation}
      \lnot a \lor \lnot\parens{b \lor c} \equiv \lnot\parens{a \land b} \land \lnot\parens{b \land c}
    \end{equation}

    Do it again (inside the parentheses).

    \begin{equation}
      \lnot a \lor \parens{\lnot b \land \lnot c} \equiv \parens{\lnot a \lor \lnot b} \land \parens{\lnot b \lor \lnot c}
    \end{equation}

    Let $p,q,r = \lnot a, \lnot b, \lnot c$, respectively.

    \begin{equation}
      p \lor \parens{q \land r} \equiv \parens{p \lor q} \land \parens{q \lor r}
    \end{equation}
  }

  \Exercise{If

    \begin{equation}
        \lnot A \implies B \implies \lnot C
    \end{equation}

    then does

    \begin{equation}
      C \Qimplies A
    \end{equation}
  }

  \Answer{Yes.

    \begin{proof}
        By transition,

      \begin{alignmath}{ccccc}
        \lnot A & \implies & B & \implies & \lnot C \\
        \lnot A &          &   & \implies & \lnot C \\
      \end{alignmath}

      By the reversal property:
      \begin{equation}
        A \impliedby C
      \end{equation}
    \end{proof}
  }

  \Exercise{If
    \begin{equation}
        A \notimplies B \implies \lnot C
    \end{equation}

    \begin{equation}
      A \Qimplies C
    \end{equation}
  }

  \Answer{Not by necessity.
    \begin{proof}
      By transition:
      \begin{alignmath}{ccccc}
        A & \notimplies & B & \implies & \lnot C \\
        A & \notimplies &   &          & \lnot C \\
      \end{alignmath}

      This does not imply

      \begin{equation}
        A \implies C
      \end{equation}
      
      To put it another way
      \begin{rclmath}
        A & \notimplies & \lnot C \\
          & \centernot\Downarrow &  \\
        A & \implies & C
      \end{rclmath}
    \end{proof}
  }

  \Exercise{If $A \implies B \impliedby C$, then

    \begin{equation}
      A \Qiff C
    \end{equation}
  }

  \Answer{No, there's nothing to indicate that $A$ and $C$ have any relation
    whatsoever.}

  \Exercise{ I'm sorry, but
    \begin{equation}
        A \land \lnot\parens{B \land \parens{C \lor D}} \Qequiv A
        \land \parens{B \lor C} \land \parens{B \lor D}
    \end{equation}
  }

  \Answer{No
    \begin{proof}
      Let's assume that it's true.
      \begin{rclmath}
        A \land \lnot\brackets{B \land \parens{C \lor D}}
          & \equiv & A \land \parens{B \lor C} \land \parens{B \lor D} \\
      \end{rclmath}
      Let $B, C, D$ all be true. Then:

      \begin{rclmath}
        A \land \lnot\brackets{\true \land \parens{\true \lor \true}}
          & \equiv & A \land \parens{\true \lor \true} \land \parens{\true \lor \true} \\
        A \land \lnot\brackets{\true \land \parens{\true \lor \true}}
          & \equiv & A \land \true \land \true \\
        A \land \lnot\brackets{\true \land \parens{\true \lor \true}}
          & \equiv & A \land \true \\
        A \land \lnot\brackets{\true \land \true}
          & \equiv & A \land \true \\
        A \land \lnot\true
          & \equiv & A \land \true \\
        A \land \lnot\true
          & \equiv & A \land \true \\
        \false
          & \equiv & \true \\
      \end{rclmath}

      Which is obviously false.
    \end{proof}
  }
\end{ExcList}

\ss{Wait, are you going to tell me what a monoid is?}

Okay, you got me!

So, monoids, like I said, are sets where you can stick the elements together to
get a new element. 

$\mset{\true,\false}$, along with the operator $\land$ is a monoid. You can take
any two elements of that set, stick them together with $\land$, and get back
another element of the set.

Additionally, $\mset{\true,\false}$, along with the operator $\lor$ forms a
monoid.

The definition of $\land$ says that if you take any boolean $a$ and $\land$ it
with $\true$, you'll get back $a$. This is called the ``monoidal identity''.

The same thing happens with $\lor$ and $\false$.

That's all a monoid is. The root ``mono'' means ``one'', so a ``monoid'' has
something to do with ``one''. It actually comes from multiplication. Numbers,
along with $\times$, form a monoid.

If you take any two numbers $x$ and $y$, $x \times y$ is also a
number. Likewise, if you take any number $x$, $x \times 1$ is just $x$. $1$ in
this case is the monoidal identity.

I bet you went through high school never hearing the term \term{monoid}! That's
criminal, given how useful the concept turns out to be, and how simple the
concept is!

This is the purpose of abstract algebra: to recognize patterns, and use the
patterns to relate two seemingly disjoint concepts. 

\begin{quote}
    Almost nothing is mutually exclusive. That is one of the usual failings of
    thinking by people, is that they think things are exclusive that aren't
    really.

    -- Larry Wall, creator of the Perl programming
    language.\cite{larry-wall-interview}
\end{quote}

Abstract algebra is all about demonstrating Larry Wall's point.