\ch{Props, proofs, and providence}
\label{ch:props}
\label{props}
\label{ch:logic}
\label{logic}

Alright, you survived the first chapter! (Or, I guess, the second
chapter).

This chapter goes over how mathematicians construct proofs. Even if you are a
Fields medalist (the math equivalent of the Nobel prize), I suggest you read
this chapter. My proof calculus (yes, it's really called that, I know\dots) is
much different than most.

With that out of the way, let's move on!

For a long time, mathematicians would deal with statements that were either
true or false. This was called \term{classical logic}. 

In the 1930s, a man named Kurt \godel\ proved a rather shocking fact: in any
non-trivial formal arithmetic system, there will always be statements that are
true, but cannot be proven. So, if we can't \xtb{prove} something, how do we
know whether or not it's true?

\godel\ was somewhat of a character. He was extremely paranoid about being
poisoned, so he refused to eat anything unless his wife prepared it for him.
Towards the end of his life, his wife was hospitalized for an extensive period
of time, during which \godel\ refused to eat, and therefore starved to
death. \cite{w-godel}

Getting back on the topic, with regard to how to decide whether or not
unprovable things are true, the answer is to not think of things in terms of
\term{true} and \term{false}, but rather in the terms of \term{provable} and
\term{unprovable}. Anything that's provable is true. Anything that isn't
provable, we don't care whether or not it's true, because we can't prove it.
\nocite{w-godel-incompleteness}

Further supporting this line of thinking is the liar's paradox:

\begin{quote}
    This statement is false.
\end{quote}

If we assume that the statement is true, then the statement makes itself
false. If you assume the statement is false, it makes itself true. This is an
example of a statement that cannot be said to be true or false. There are
similar statements in math, were we can't even say it's true or false.

In the case of the liar's paradox, we have a pretty quick cop-out, by just
saying that it's not provable. We don't care whether or not it's true.

For this, we have two values in math, which we throw around a lot: $\top$, and
$\bot$. The symbols are called \term{top} and \term{bottom}, respectively. They
refer, respectively to `provable' and `unprovable'.

\xtb{Why don't you just call the symbols `provable' and `unprovable'?}

It's sort of like how $+$ and $-$ refer to the concepts addition and
subtraction, but are pronounced `plus' and `minus'.

You're probably still used to calling propositions `true' and `false'. That's
okay. Over time, you'll get used to calling them `top' and `bottom'. Even
mathematicians will accidentally call them `true' and `false'.

\input{3/1-if-then.ltx}
\newpage
\input{3/2-walk.ltx}
\newpage
